Python as a machine learning language: experiences from the scikit-learn project

### 2013-04-27 api.tex
typo

% TODO also discuss scipy.sparse
Input to an estimator is typically given as NumPy arrays \citep{vanderwalt2011};
typically, a clustering or dimensionality reduction algorithm
takes a single NumPy array, conventionally called $X$ of size $n \times d$
to represent $n$ samples of $d$ features each.
A regression or classification estimator takes, at training time,
an additional array of size $n$, by convention called $y$,
that represents the target variable.
While the NumPy array may seem like a ``bare'' representation of data
when compared to more object-oriented representations
(again, Weka is an example of this style of API),
it brings the prime advantages of allowing \textit{vector operations}
on entire arrays,
which are often orders of magnitude faster than the corresponding Python loops
while at the same time keeping the code short and readable.
Although conversion to the NumPy format may seem like a burden on the user,
we argue that this is not the case.
Python's lists (its dynamic array structure)
are easily transformed into NumPy arrays
and many scientific users of Python will already be handling such arrays,
since they pervade in other science-oriented Python packages.
For tasks where input is likely to consists of text files
or semi-structured objects, ``vectorizers'' are provided
that efficiently convert from such data formats to the NumPy format.
Finally, offering a custom datatype for samples
would require a conversion \textit{in any case}.

>>>>

% TODO also discuss scipy.sparse
Input to an estimator is typically given as NumPy arrays \citep{vanderwalt2011};
typically, a clustering or dimensionality reduction algorithm
takes a single NumPy array, conventionally called $X$ of size $n \times d$
to represent $n$ samples of $d$ features each.
A regression or classification estimator takes, at training time,
an additional array of size $n$, by convention called $y$,
that represents the target variable.
While the NumPy array may seem like a ``bare'' representation of data
when compared to more object-oriented representations
(again, Weka is an example of this style of API),
it brings the prime advantages of allowing \textit{vector operations}
on entire arrays,
which are often orders of magnitude faster than the corresponding Python loops
while at the same time keeping the code short and readable.
Although conversion to the NumPy format may seem like a burden on the user,
we argue that this is not the case.
Python's lists (its dynamic array structure)
are easily transformed into NumPy arrays
and many scientific users of Python will already be handling such arrays,
since they are pervasive in other science-oriented Python packages.
For tasks where input is likely to consists of text files
or semi-structured objects, ``vectorizers'' are provided
that efficiently convert from such data formats to the NumPy format.
Finally, offering a custom datatype for samples
would require a conversion \textit{in any case}.
